---
title: "Homework3"
output:
  pdf_document: default
  html_document: default
date: "2023-02-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

```{r}
## Importing the libraries
library(readr)
library(ggplot2)
library(dplyr)
library(tidyverse)
```

```{r}
## Importing data from Enrollment.csv and replacing all the reserve codes as missing values.

df1data = read_csv("Enrollment.csv",na = c(-9,-3,-5,-6,-8,-11,"NA"))
```

```{r}
head(df1data)
```

```{r}
## Selecting the required columns
df1datacols <- select(df1data, 
                      LEAID,
                      SCHID, 
                      SCH_ENR_HI_M,
                      SCH_ENR_HI_F,
                      SCH_ENR_AM_M,
                      SCH_ENR_AM_F, 
                      SCH_ENR_AS_M,
                      SCH_ENR_AS_F,
                      SCH_ENR_HP_M,
                      SCH_ENR_HP_F,
                      SCH_ENR_BL_M, 
                      SCH_ENR_BL_F,
                      SCH_ENR_WH_M, 
                      SCH_ENR_WH_F,
                      SCH_ENR_TR_M,
                      SCH_ENR_TR_F)
head(df1datacols)
```

```{r}
## Using pivot longer to transform data.
df1datatransform <- 
  df1datacols %>% pivot_longer(cols=c(`SCH_ENR_HI_M`, 
                                      `SCH_ENR_HI_F`, 
                                      `SCH_ENR_AM_M`, 
                                      `SCH_ENR_AM_F`, 
                                      `SCH_ENR_AS_M`, 
                                      `SCH_ENR_AS_F`, 
                                      `SCH_ENR_HP_M`, 
                                      `SCH_ENR_HP_F`, 
                                      `SCH_ENR_BL_M`, 
                                      `SCH_ENR_BL_F`, 
                                      `SCH_ENR_WH_M`,
                                      `SCH_ENR_WH_F`, 
                                      `SCH_ENR_TR_M`, 
                                      `SCH_ENR_TR_F`), 
                               names_to = 'Race Gender',
                               values_to ='Enrollment Count')

head(df1datatransform)
```

```{r}
## Renaming for better readability.

df1datatransform <-
  rename(df1datatransform, `School ID` = SCHID, `Lea ID` = LEAID)
df1datarename <- 
  df1datatransform %>% 
  mutate(`Race Gender` = recode(`Race Gender`,
                                "SCH_ENR_HI_M" = "HisanicLatino Male", 
                                "SCH_ENR_HI_F"= "HisanicLatino Female", 
                                "SCH_ENR_AM_M" = "AIndianANative Male" ,
                                "SCH_ENR_AM_F" = "AIndianANative Female",
                                "SCH_ENR_AS_M" = "Asian Male" ,
                                "SCH_ENR_AS_F" = "Asian Female",
                                "SCH_ENR_HP_M" = "NativeHawaiian Male", 
                                "SCH_ENR_HP_F" = "NativeHawaiian Female", 
                                "SCH_ENR_BL_M" = "AfricanAmerican Male", 
                                "SCH_ENR_BL_F" = "AfricanAmerican Female",
                                "SCH_ENR_WH_M" = "White Male" ,
                                "SCH_ENR_WH_F" = "White Female",
                                "SCH_ENR_TR_M" = "TwoRace Male",
                                "SCH_ENR_TR_F" = "TwoRace Female"))

head(df1datarename)
```

```{r}
## Calculating the total enrollment count.

class(df1datarename$`Enrollment Count`)
totalEnrollment <- sum(df1datarename[,'Enrollment Count'], na.rm = TRUE)
totalEnrollment
```

```{r}
## Aggregating data.
df1dataagg <- 
  aggregate(df1datarename$`Enrollment Count`, 
            list(df1datarename$`Race Gender`),sum, na.rm=TRUE)
head(df1dataagg)
```
```{r}
## Calculating and add proportion column. 
df1datafinal <-
  df1dataagg %>% mutate(proportion = df1dataagg$x / totalEnrollment)
head(df1datafinal)
```


```{r}
## Plotting the graph for Problem 1.
ggplot(df1datafinal, aes(x=`Group.1`, y= proportion)) +
geom_bar(stat = "identity", fill="lightblue") +
labs(title="Proportion of enrolled students of every race and gender out 
the total students across all schools",
x="Race and Gender", y="Proportion") +
coord_flip() +
theme_minimal()
```

## Observations:

From the above plot, we can see that the proportion of male as well as female
"Native Hawaiian or Other Pacific Islander" is the lowest out of the total
number of students across all school.

We can also observe that the proportion of male and female "American Indian or 
Alaska Native" is the second lowest.

Overall for each Race-Gender combination, we can observe that the proportion of
males is higher as compared to females.


## Problem 2

```{r}
## Importing data.
df2data = read_csv("Advanced Placement.csv",na = c(-9,-3,-5,-6,-8,-11,"NA"))
head(df2data)
```

```{r}
## Selecting the required columns.
df2datacols <-df2data %>%
  select(LEAID,SCHID,
         SCH_APENR_HI_M,
         SCH_APENR_HI_F,
         SCH_APENR_AM_M,
         SCH_APENR_AM_F,
         SCH_APENR_AS_M,
         SCH_APENR_AS_F, 
         SCH_APENR_HP_M,
         SCH_APENR_HP_F,
         SCH_APENR_BL_M,
         SCH_APENR_BL_F,
         SCH_APENR_WH_M,
         SCH_APENR_WH_F,
         SCH_APENR_TR_M,
         SCH_APENR_TR_F) %>%
  filter(df2data$SCH_APENR_IND == 'Yes')

head(df2datacols)
```

```{r}
## Using pivot_longer to transform data.
df2datatransform <-
  df2datacols %>% pivot_longer(cols=c(colnames(df2datacols)[-c(1,2)]),
                               names_to = 'Race Gender',
                               values_to ='Enrollment Count')
head(df2datatransform)
```

```{r}
## Renaming for better readability.
df2datatransform <-rename(df2datatransform, `School ID` = SCHID, `Lea ID` = LEAID)
df2datarename <- 
  df2datatransform %>%
  mutate(`Race Gender` =
           recode(`Race Gender`,
                  "SCH_APENR_HI_M" = "HisanicLatino Male",
                  "SCH_APENR_HI_F"= "HisanicLatino Female", 
                  "SCH_APENR_AM_M" = "AIndianANative Male" ,
                  "SCH_APENR_AM_F" = "AIndianANative Female", 
                  "SCH_APENR_AS_M" = "Asian Male" , 
                  "SCH_APENR_AS_F" = "Asian Female", 
                  "SCH_APENR_HP_M" = "NativeHawaiian Male", 
                  "SCH_APENR_HP_F" = "NativeHawaiian Female",
                  "SCH_APENR_BL_M" = "AfricanAmerican Male",
                  "SCH_APENR_BL_F" = "AfricanAmerican Female", 
                  "SCH_APENR_WH_M" = "White Male" , 
                  "SCH_APENR_WH_F" = "White Female", 
                  "SCH_APENR_TR_M" = "TwoRace Male", 
                  "SCH_APENR_TR_F" = "TwoRace Female"))

head(df2datarename)
```

```{r}
## Calculating the total enrollment.
totalEnrollment <- sum(df2datarename[,'Enrollment Count'], na.rm = TRUE)
totalEnrollment
```


```{r}
## Aggregating the data.
df2dataagg <- 
  aggregate(df2datarename$`Enrollment Count`, 
            list(df2datarename$`Race Gender`),sum, na.rm=TRUE) 
head(df2dataagg)
```

```{r}
## Adding the proportion column.
df2datafinal<- 
  df2dataagg %>% mutate(proportion = df2dataagg$x / totalEnrollment)
head(df2datafinal)
```

```{r}
## Plotting graph for Problem 2.
ggplot(df2datafinal, aes(x=`Group.1`, y= proportion)) +
geom_bar(stat = "identity" , fill="lightpink") +
labs(title="Proportion of AP students for every race and gender out the
total AP students across all schools",
x="Race and Gender", y="Proportion") +
coord_flip() +
theme_minimal()
```

## Observations:
From the above plot, we can see that the proportion of male as well as female
"Native Hawaiian or Other Pacific Islander" students in AP class is the lowest 
out of the total number of AP students across all school. 

We can also observe that the proportion of male and female "American Indian or
Alaska Native" students in AP class is the second lowest. 

Overall for each Race-Gender combination, we can observe that the proportion of 
females students is higher as compared to male students.



## Problem 3

```{r}
## Making copies of data imported for previous problems.
df3dataenrollment <- df1datacols
df3dataap <- df2datacols

head(df3dataenrollment)
head(df3dataap)
```

```{r}
## Adding proportion of students of color out of all enrolled students at the school.
df3dataenrollment$color <-
  rowSums(df3dataenrollment[,c(names(df3dataenrollment)[3:12],
                               names(df3dataenrollment)[15:16])])/rowSums(
                                 df3dataenrollment[,c(names(df3dataenrollment)
                                                      [3:16])])
df3dataenrollment
```

```{r}
## Adding the proportion of students of color in at least one AP class out of all students in at least one AP class. 

df3dataap$apcolor <- 
  rowSums(df3dataap[,c(names(df3dataap)[3:12],
                       names(df3dataap)[15:16])])/rowSums(
                         df3dataap[,c(names(df3dataap)[3:16])])
head(df3dataap)
```

```{r}
## Joining both the tables on SCHID and LEAID as by combining both we get unique enteries.
df3final <- merge(df3dataenrollment,df3dataap,by =c("SCHID","LEAID"))
head(df3final)
```
```{r}
## Plotting graph for problem 3
ggplot(df3final, aes(x=`color`, y= `apcolor`)) +
geom_point() + geom_abline(slope = 1,  color = "yellow") +
labs(title="Proportion of students of color out of total enrolled students at the
school vs the proportion of students of color in AP class",
x="Enrollment Proportion of students of color", y="AP Proportion of students of color") +
theme_minimal()
```
## Observations:

From the above graph we can observe that as the enrollment proportion of students
of color increase, the proportion of students of color in AP class also increases. 

## Problem 4

```{r}
library(RSQLite)
```

```{r}
con <- dbConnect(SQLite(), "dblp.db")
```

```{r}
dbListTables(con)
```

```{r}
## Querying general table.
datageneral <- dbGetQuery(con,
           "SELECT *
           FROM general")
head(datageneral)
```

```{r}
## Querying authors table.
dataauthor <- dbGetQuery(con,
           "SELECT *
           FROM authors")
head(dataauthor)
```

```{r}
## Joining both the tables.
datasol4<-dbGetQuery(con,
           "SELECT DISTINCT name, year, gender, prob
           FROM general JOIN authors using(k) WHERE gender IN 
           ('M','F') AND prob >= 0.90")
head(datasol4)
```


```{r}
## Plot for Problem 4.
ggplot(datasol4, aes(x=year)) +
geom_bar() + 
facet_wrap(~gender) +
labs(title="Total number of distinct male and female authors
published each year", x="Year", y="Count") +
theme_minimal()
```
## Observations:

From the above graph we can observe that the number of overall authors is 
increasing over the years. A comparison with respect to gender shows that on 
average there are more number of male authors than female authors published
each year. 

## Problem 5

```{r}
## Plot for Problem 5.
ggplot(datasol4, aes(x=year, fill = gender)) +
geom_bar(position = "fill") +
scale_fill_discrete(name="Gender") +
labs(title="Proportions of distinct male authors vs distinct female
authors published each year",
x="Year", y="Proportion") +
theme_minimal()
```


## Observation

From the above graph, we observe that for the years 1960-1964, there are no 
female authors published. Around 1970s the proportion of female authors 
published starts increasing over the years. Overall, female authors published 
each year are less as compared to the male authors.
