---
title: "Homework 4"
output:
  pdf_document: default
  html_document: default
date: "2023-03-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Loading the required libraries.
library(ggplot2)
library(dplyr)
library(modelr)
```

```{r}
# Loading the data.
load(file = '37938-0001-Data.rda')
df<-da37938.0001
```

```{r}
# Subset the required columns.
dfdata1 <- df %>% select(RACE,WEIGHT) 
head(dfdata1)
```
```{r}
## Replacing the numbering in RACE column with space 
dfdata1$RACE <- gsub("^\\(\\d+\\)\\s+", "", dfdata1$RACE)
head(dfdata1)
```
```{r}
# Calculating the proportion using un-weighted column.
freqTable <- table(dfdata1$RACE)
totalCount <- sum(freqTable)
unweightedProportion = freqTable/totalCount
dfdata1final<-data.frame(values = unweightedProportion )
dfdata1final
```

```{r}
# Plotting the graph for un-weighted proportion of trans people of each race/ethnicity.
ggplot(dfdata1final, aes(x=values.Var1, y = values.Freq)) +
geom_bar(stat = "identity" , fill="lightpink") +
labs(title="Unweighted proportions of trans people of each Race/Ethnicity",
x="Race", y="Unweighted Proportions") +
coord_flip() +
theme_minimal()
```
```{r}
# Calculating the proportion using weighted column.
dfdata1weight <- dfdata1 %>% group_by(RACE) %>% 
  summarise(weights = sum(WEIGHT))%>%
  mutate(weightedproportion = weights / sum(weights))
dfdata1weight
```

```{r}
# Plotting the graph for weighted proportion of trans people of each race/ethnicity.
ggplot(dfdata1weight, aes(x=RACE, y= weightedproportion)) +
geom_bar(stat = "identity" , fill="lightblue") +
labs(title="Weighted proportions of trans people of each Race/Ethnicity",
x="Race", y="Weighted Proportions")  +
coord_flip() +
theme_minimal()
```

## Observations:

From the above graphs, we can observe that in both weighted and un-weighted
graphs the proportion of trans people of White race/ethnicity is the highest. 

In the graph for "Un-weighted proportion of trans people of each race and ethnicity", 
Middle Eastern race/ethnicity has the lowest proportion, however in the 
"Weighted proportion of trans people of each race and ethnicity", American Indian
is the least.

On comparing both the graphs, we observe that the proportion of White race/ethnicity
has decreased from 0.68 for un-weighted to 0.56 for weighted and similarly the
proportion of American Indian decreased from 0.007 for un-weighted to 0.003
for weighted which means both these races were over-represented in the survey 
sample compared to the population. Other than this, population of all other 
races have increased. 

## Problem 2

```{r}
## Subset the required columns
dfdata2 <- df %>% select(SEXUALID,WEIGHT)
head(dfdata2)
```

```{r}
## Replacing the numbering in SEXUALID column with space 
dfdata2$SEXUALID <- gsub("^\\(\\d+\\)\\s+", "", dfdata2$SEXUALID)
head(dfdata2)
```

```{r}
# Calculating the proportion using un-weighted column.
freqTable <- table(dfdata2$SEXUALID)
totalCount <- sum(freqTable)
unweightedProportion = freqTable/totalCount
dfdata2final<-data.frame(values = unweightedProportion )
dfdata2final
```

```{r}
# Plotting the graph for un-weighted proportion of trans people of each sexual orientation.
ggplot(dfdata2final, aes(x=values.Var1, y = values.Freq)) +
geom_bar(stat = "identity" , fill="lightpink") +
labs(title="Unweighted proportions of trans people of each sexual
orientation",
x="Sexual Orientation", y="Unweighted Proportions") +
coord_flip() +
theme_minimal()
```

```{r}
# Calculating the proportion using weighted column.
dfdata2weight <- dfdata2 %>% group_by(SEXUALID) %>% 
filter(!is.na(SEXUALID))  %>%
summarise(weights = sum(WEIGHT))%>%
mutate(weightedproportion = weights / sum(dfdata2$WEIGHT))
dfdata2weight
```

```{r}
# Plotting the graph for weighted proportion of trans people of each sexual orientation.
ggplot(dfdata2weight, aes(x=SEXUALID, y= weightedproportion)) +
geom_bar(stat = "identity" , fill="lightblue") +
labs(title="Weighted proportions of trans people of each sexual orientation",
x="Sexual Orientation", y="Weighted Proportions") +
coord_flip() +
theme_minimal()
```

## Observation:

From the above graphs, we can observe that in the un-weighted graphs the proportion 
of trans people of Straight/Heterosexual sexual orientation is the highest whereas
in the weighted graph, the proportion of Bisexual orientation is the highest. 

In the graph for "Un-weighted proportion of trans people of each sexual orientation", 
Same-Gender loving and asexual spectrum  have a lower representation, however in the 
"Weighted proportion of trans people of each sexual orientation", Same-Gender loving
has the lowest proportion.

On comparing both the graphs, the proportion of Asexual spectrum and Pansexual 
orientation show significant increase from un-weighted to weighted however
the proportion of Straight/Heterosexual sexual orientation add same-gender loving
decreases from un-weighted to weighted.

## PROBLEM 3

Graph 1 :

```{r}
# Visualization of life satisfaction versus Social well-being.
ggplot(df, aes(x=(SOCIALWB_I), y=(LIFESAT_I))) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method="lm", color="red") +
  labs(x="Social well-being", y="Satisfaction with life") +
  theme_minimal()

ggplot(df, aes(x=log(SOCIALWB_I), y=log(LIFESAT_I))) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method="lm", color="red") +
  labs(x="Social well-being", y="Satisfaction with life") +
  theme_minimal()
```

## Observation:

The above relationship between "Satisfaction with life" and "Social well-being"
has a positive correlation and graph is mostly linear. The data is skewed towards 
left and performing log-transformation also does not help and hence we cannot 
use this as an ideal predictor.

Graph 2 :

```{r}
# Visualization of life satisfaction versus Non-affirmation of gender identity.
ggplot(df, aes(x=(NONAFFIRM_I), y=(LIFESAT_I))) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method="lm", color="red") +
  labs(x="Non-affirmation of gender identity", y="Satisfaction with life") +
  theme_minimal()

ggplot(df, aes(x=log2(NONAFFIRM_I), y=log2(LIFESAT_I))) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method="lm", color="red") +
  labs(x="Non-affirmation of gender identity", y="Satisfaction with life") +
  theme_minimal()
```

## Observation:

The above relationship between "Satisfaction with life" and "Non-affirmation of 
gender identity" has a negative correlation and the graph does not look mostly 
linear as there a big dip. Performing log-transformation also does not affect the
relationship and makes the data skewed and hence we cannot use this as an ideal
predictor.

Graph 3:

```{r}
# Visualization of life satisfaction versus Non-disclosure of gender identity.
ggplot(df, aes(x=(NONDISCLOSURE_I), y=(LIFESAT_I), na.rm = TRUE)) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method="lm", color="red") +
  labs(x="Non-disclosure of gender identity", y="Satisfaction with life") +
  theme_minimal()

ggplot(df, aes(x=log2(NONDISCLOSURE_I), y=log2(LIFESAT_I), na.rm = TRUE)) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method="lm", color="red") +
  labs(x="Non-disclosure of gender identity", y="Satisfaction with life") +
  theme_minimal()
```

## Observation:

The above relationship between "Satisfaction with life" and "Non-disclosure 
of gender identity" has a negative correlation and does not look linear as we
observe a few dips. Performing log-transformation also does not affect the 
relationship and hence this cannot be considered as an ideal predictor for modelling.

Graph 4:

```{r}
# Visualization of life satisfaction versus Healthcare stereotype threat.

ggplot(df, aes(x=(HCTHREAT_I), y=(LIFESAT_I), na.rm = TRUE)) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method="lm", color="red") +
  labs(x="Healthcare stereotype threat", y="Satisfaction with life") +
  theme_minimal()

ggplot(df, aes(x=log2(HCTHREAT_I), y=log2(LIFESAT_I), na.rm = TRUE)) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method="lm", color="red") +
  labs(x="Healthcare stereotype threat", y="Satisfaction with life") +
  theme_minimal()

```

## Observation:

The above relationship between "Satisfaction with life" and "Healthcare 
stereotype threat" has a string negative correlation and looks linear. 
As the data is not skewed performing log transformation does not make the 
relationship any more linear than the original graph. Hence this can be an 
ideal candidate for modelling.

Graph 5:

```{r}
# Visualization of life satisfaction versus Mental distress/disorder.
ggplot(df, aes(x=(KESSLER6_I), y=(LIFESAT_I), na.rm = TRUE)) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method="lm", color="red") +
  labs(x="Mental distress/disorder", y="Satisfaction with life") +
  theme_minimal()

ggplot(df, aes(x=log(KESSLER6_I), y=log(LIFESAT_I), na.rm = TRUE)) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method="lm", color="red") +
  labs(x="Mental distress/disorder", y="Satisfaction with life") +
  theme_minimal()
```

## Observation:

The above relationship between "Satisfaction with life" and "Mental 
distress/disorder" has strong negative correlation and looks look linear. 
As the data is mostly linearly distributed across the x axis, performing log 
transformation does not improve the relationship. This can be treated as an ideal
predictor for modelling. 

Graph 6:

```{r}
# Visualization of life satisfaction versus Mental distress/disorder.
ggplot(df, aes(x=(EVERYDAY_I), y=(LIFESAT_I), na.rm = TRUE)) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method="lm", color="red") +
  labs(x="Everyday discrimination", y="Satisfaction with life") +
  theme_minimal()

ggplot(df, aes(x=log(EVERYDAY_I), y=log(LIFESAT_I), na.rm = TRUE)) +
  geom_point() +
  geom_smooth() +
  geom_smooth(method="lm", color="red") +
  labs(x="Everyday discrimination", y="Satisfaction with life") +
  theme_minimal()
```
## Observation:

The above relationship between "Satisfaction with life" and "Everyday discrimination" 
has a negative correlation and we can see that the data is skewed. Performing 
log-transformation also does not affect the relationship and hence this would
not be an ideal predictor for modelling. 


## Problem 4

# Justification:

From the above satisfaction vs the candidate predictors plots,
I decided to go with the Mental distress/disorder as the predictor for Problem 4.
After plotting the graphs, Healthcare stereotype threat as well 
as Mental distress/disorder plots show a linear relationship and hence should be
suitable to be used for the linear regression model. 
Using Mental distress/disorder gives a linear qq-plot which further
helps verify that this can be treated as an ideal predictor.

```{r}
# Linear regression model for life satisfaction using a single predictor:
fit1 <- lm((LIFESAT_I) ~ (KESSLER6_I), data=df)
summary(fit1)
```

```{r}
# Plotting the predictions.
df %>%
  add_predictions(fit1) %>%
  ggplot(aes(x=KESSLER6_I)) +
  geom_point(aes(y=LIFESAT_I)) +
  geom_line(aes(y=pred), color="red", size=1) +
  coord_cartesian(xlim=range(df$KESSLER6_I),
                  ylim=range(df$LIFESAT_I)) +
  labs(x="Mental distress/disorder", y="Satisfaction with life") +
  theme_minimal()
```

```{r}
# Plotting the residuals
df %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(x=(KESSLER6_I))) +
  geom_point(aes(y=resid)) +
  labs(x="Mental distress/disorder", y="Residuals") +
  theme_minimal()
```
```{r}
# Normality of residuals.
df %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() +
  theme_minimal()

df %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(x=resid)) +
  geom_histogram(bins=100) +
  labs(x="Residuals") +
  theme_minimal()
```

## Observations:

From the above graphs, the qq-plot looks approximately linear. There are a few
outliers, however that does not change the linearity of the graph and can be 
ignored. 

## Problem 5

# Plotting residulas with other candidate predictors:

```{r}
# Plotting the residuals
df %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(x=(SOCIALWB_I))) +
  geom_point(aes(y=resid)) +
  labs(x="Social well-being", y="Residuals") +
  theme_minimal()

df %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(x=(NONAFFIRM_I))) +
  geom_point(aes(y=resid)) +
  labs(x="Non-affirmation of gender identity", y="Residuals") +
  theme_minimal()

df %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(x=(NONDISCLOSURE_I))) +
  geom_point(aes(y=resid)) +
  labs(x="Non-disclosure of gender identity", y="Residuals") +
  theme_minimal()

df %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(x=(HCTHREAT_I))) +
  geom_point(aes(y=resid)) +
  labs(x="Healthcare stereotype threat", y="Residuals") +
  theme_minimal()

df %>%
  add_residuals(fit1, "resid") %>%
  ggplot(aes(x=(EVERYDAY_I))) +
  geom_point(aes(y=resid)) +
  labs(x="Everyday discrimination", y="Residuals") +
  theme_minimal()

```
## Observation:

From the above residual plots, it can be seen that Healthcare stereotype threat
and Non-affirmation of gender identity have the most random scatter and can be
added to the model. In this case, I will be using Healthcare stereotype threat 
as that generates the most linear QQ plot.

```{r}
# Linear regression model for life satisfaction using a single predictor:
fit2 <- lm((LIFESAT_I) ~ (KESSLER6_I) + HCTHREAT_I  , data=df)
summary(fit2)
```

```{r}
# Normality of residuals.
df %>%
  add_residuals(fit2, "resid") %>%
  ggplot(aes(sample=resid)) +
  geom_qq() +
  theme_minimal()

df %>%
  add_residuals(fit2, "resid") %>%
  ggplot(aes(x=resid)) +
  geom_histogram(bins=100) +
  labs(x="Residuals") +
  theme_minimal()
```

## Observation:

From the above residual plots, it can be seen that both the predictors combined
create a linear QQ plot.
